Inference for Trans-dimensional Bayesian Models with Diffusive Nested Sampling

### 2014-08-27 Documents/Paper/ms.tex
More references and writing

In other studies, it has been common to do separate runs of Nested Sampling
(or another method),
with $N$ fixed at various trial values \citep[e.g.][]{fengji}.
Then, the posterior for $N$ can be calculated
based on the estimates of the evidence obtained as a function of $N$. However,
in this paper, we use trans-dimensional MCMC moves to infer $N$. In other words,
$N$ is treated as just another parameter, and the marginal likelihood defined
in Equation~\ref{eq:evidence}, which we can compute, also includes the sum over
the unknown parameter $N$. Our motivation
for using Nested Sampling is not primarily to calculate $\mathcal{Z}$.
The motivation for using Nested Sampling, rather than just sampling the
posterior using trans-dimensional MCMC, is that the posterior often contains
difficult features, such as strong dependencies or multiple modes, that cause
problems with mixing. Diffusive Nested Sampling
(DNS) replaces the posterior distribution with an alternative target
distribution composed of a mixture of the prior distribution with other, more
constrained distributions, facilitating mixing between the multiple modes
and along degeneracy curves.
This is similar to the idea of simulated annealing,
where the target distribution is modified from the posterior to something
``easier'', however,
the sequence of distributions used in Nested Sampling avoids some of the
problems with simulated annealing: in particular, there is no need to choose
a temperature schedule, and the method does not fail when phase transitions
are present.

Since DNS is effectively the Metropolis-Hastings algorithm applied to a
distribution other than the posterior, proposal distributions are needed.
In Section~\ref{sec:proposals}, we outline a set of proposal distributions used
to explore the hypothesis space of possible values for $N$, $\hyperparams$,
and $\{\xx_i\}$.

@@

\begin{figure}
\begin{center}
\includegraphics[scale=0.5]{galaxyfield_likelihood.pdf}
\caption{{\bf Top panel: }
The shape of the likelihood function with respect to prior mass for the
GalaxyField data. Note the phase
transition at $\log(X) \approx -100$ nats. This separates models with just
the few brightest galaxies from models with many faint galaxies as well.
While this doesn't affect the
posterior distribution (unlike the sinewave example), it would affect calculation
of the marginal likelihood if annealing were used.
{\bf Bottom panel: }The posterior weights of the saved samples.
\label{fig:sinewaves_likelihood}}
\end{center}
\end{figure}

@@

@@

>>>>

Some authors have addressed these kinds of problems by
doing separate runs of Nested Sampling
(or another method),
with $N$ fixed at various trial values \citep[e.g.][]{fengji, feroz}.
Then, the posterior for $N$ can be calculated
based on the estimates of the evidence obtained as a function of $N$. However,
in this paper, we use trans-dimensional MCMC moves to infer $N$. In other words,
$N$ is treated as just another parameter, and the marginal likelihood defined
in Equation~\ref{eq:evidence}, which we can compute, also includes the sum over
the unknown parameter $N$. Our motivation
for using Nested Sampling is not primarily to calculate $\mathcal{Z}$.
The motivation for using Nested Sampling, rather than just sampling the
posterior using trans-dimensional MCMC, is that the posterior often contains
difficult features, such as strong dependencies or multiple modes, that cause
problems with mixing. Diffusive Nested Sampling
(DNS) replaces the posterior distribution with an alternative target
distribution composed of a mixture of the prior distribution with other, more
constrained versions of the prior, facilitating mixing between the multiple modes
and along degeneracy curves.
This is similar to the idea of annealed importance sampling
\citep{neal} or parallel tempering \citep{pt},
where the target distribution is modified from the posterior to something
``easier''. However,
the sequence of distributions used in Nested Sampling avoids some of the
problems with annealing-based methods. In particular, there is no need to choose
a temperature schedule (a potentially large set of tuning parameters),
and the method does not fail when phase transitions
are present.

Since DNS is effectively the Metropolis-Hastings algorithm applied to a
distribution other than the posterior, proposal distributions are needed
which define how a walker moves around the hypothesis space.
In Section~\ref{sec:proposals}, we outline a set of proposal distributions used
to explore the hypothesis space of possible values for $N$, $\hyperparams$,
and $\{\xx_i\}$.

@@

\begin{figure}
\begin{center}
\includegraphics[scale=0.5]{galaxyfield_likelihood.pdf}
\caption{{\bf Top panel: }
The shape of the likelihood function with respect to prior mass for the
GalaxyField data. Note the phase
transition at $\log(X) \approx -100$ nats. This separates models with just
the few brightest galaxies from models with many faint galaxies as well.
While this doesn't affect the
posterior distribution (unlike the sinewave example), it would affect calculation
of the marginal likelihood if annealing were used.
{\bf Bottom panel: }The posterior weights of the saved samples.
\label{fig:galaxyfield_likelihood}}
\end{center}
\end{figure}

@@

\bibitem[Feroz and Hobson(2014)]{feroz} Feroz, F., Hobson, 
M.~P.\ 2014.\ Bayesian analysis of radial velocity data of GJ667C with 
correlated noise: evidence for only two planets.\ Monthly Notices of the 
Royal Astronomical Society 437, 3540-3549.

@@

\bibitem[Hansmann(1997)]{pt} Hansmann, Ulrich HE., 1997, ``Parallel tempering algorithm for conformational studies of biological molecules.'', Chemical Physics Letters 281, no. 1 (1997): 140-150.
